---
title: "How well do you lift?"
author: "Mathieu C."
date: "16 décembre 2019"
output: html_document
---
## Synopsis

The goal of the following procedures is to make a model that reliably classifies (potentially new) data within five groups, labelled with letters A to E.  
  
The letters assess the quality of an exercise performed while wearing different sensors that produce the X, Y, and Z axis acceleration values as well as other values.
  
More details can be found by following this [link](https://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har), under the section *Weight Lifting Exercices Dataset* (Credits; *Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.*).  



## Data Aquisition and Manipulation

The code under details all manipulations done to the datasets (training and testing).  

```{r, warning=FALSE, message=FALSE}
# Loading libraries and setting the seed
library(caret);library(ggplot2); library(randomForest); library(gbm)
set.seed(1234)
# Downloading the files containing the training/testing sets
train.path <- 
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.path <- 
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if (!file.exists("train.csv") | !file.exists("test.csv")){
        download.file(train.path, destfile = "train.csv")
        download.file(test.path,destfile = "test.csv")
}
raw.training.set <- read.csv("train.csv", header = T, 
                             na.strings = c("#DIV/0!","NA"),as.is = T)
raw.testing.set <- read.csv("test.csv", header = T, 
                            na.strings = c("#DIV/0!","NA"), as.is = T)

# Removing columns containing NA values
na.vals <- apply(raw.training.set, 2, is.na)
na.vals.sum <- apply(na.vals, 2, sum)
na.vals.cols <- na.vals.sum > 0
training <- raw.training.set[,-which(na.vals.cols)]
testing <- raw.testing.set[,-which(na.vals.cols)]

## Further removing irrelevant columns like name, X, timestamps or ID of the serie
training <- training[,8:60]; testing <- testing[,8:60]

# Checking for high correlation in the remaining variables
m.corr <- abs(cor(training[,-53]))
diag(m.corr) <- 0

# Crearing a validation set representing 40% of the training set
intrain <- createDataPartition(training$classe, p = .60, list = F)
validation <- training[-intrain,]
training <- training[intrain,]

```

# Model Building and testing
  
We'll start with a randomForest model:

```{r, cache=TRUE}
## Building model
training$classe <- factor(training$classe)
mod.rf <- randomForest(classe~., training)

## Accuracy test
accur.df <- data.frame(predict = predict(mod.rf, newdata = validation), 
                       true = validation$classe)
accur.table <- table(accur.df)
accuracy.rf <- sum(predict(mod.rf, newdata = validation) == validation$classe)/
        length(validation$classe) ; names(accuracy.rf) <- "Random Forest Accuracy"
print(accuracy.rf); print(accur.table)
```

This is a pretty good result on the first try, random forest seems to make a great job with default settings and gives us an accuracy above 99% with a few misclassification evenly distributed among groups.  

However, we'll try different methods to see if there's an improvement.


```{r, cache=TRUE}
# Building GBM Prediction
mod.gbm <- gbm(classe~., data = training, n.trees = 500)
predict.gbm <- predict(mod.gbm, newdata = validation, n.trees = 500, type = "response")

predict.gbm.fac <- factor(apply(predict.gbm, 1, which.max))
levels(predict.gbm.fac) <- c("A", "B", "C", "D", "E")
accuracy.gbm <- sum(predict.gbm.fac == validation$classe)/ length(validation$classe) ; names(accuracy.gbm) <- "GBM Accuracy"
accur.gbm.df <- data.frame(prediction = predict.gbm.fac, truth = validation$classe)
print(accuracy.gbm); print(table(accur.gbm.df))
```
  
We can see that the acuracy is still high but lower, and the misclassification is less controlled.  
We will probably stick with the randomForest model. Stacking the models (like demonstrated below) does not work great using the "gam" method. Probably because of the very high correlation of the two models.

```{r}
# Agreement accuracy
agr.acc <- sum(predict.gbm.fac == predict(mod.rf, newdata = validation, type = "response"))/length(validation$classe)
names(agr.acc) <- "Agreement Accuracy"
print(agr.acc)
```
  
Below is the (failed) attempt to stack both models.

```{r, warning=FALSE}
# Trying to stack models
stack.df <- data.frame(pred.rf = as.numeric(predict(mod.rf, 
                                                    newdata = validation, 
                                                    type = "response")),
                       pred.gbm = as.numeric(predict.gbm.fac),
                       truth = validation$classe)
stack.cor <- cor(as.numeric(stack.df$pred.rf), as.numeric(stack.df$pred.gbm))
names(stack.cor) <- "Correlation between the two models"
mod.stack <- train(truth~., data = stack.df, method = "gam")
accuracy.stack <- sum(predict(mod.stack, stack.df)==stack.df$truth)/
  length(stack.df$truth); names(accuracy.stack) <- "Stacked Models Accuracy"
errors.mod.stack <- predict(mod.stack, stack.df) != stack.df$truth
accur.stack <- data.frame(prediction = predict(mod.stack),
                          truth = stack.df$truth)
table.stack <- table(accur.stack)
print(accuracy.stack); print(table.stack); print(stack.cor)
```
  
## Applying on the test set
  
```{r}
final.pred <- predict(mod.rf, newdata = testing)
print(final.pred)
```


## Conclusion
  
The randomForest model gives expected results and has a good interpretability, if needed.